# Transform Tables

***

```{r echo = FALSE, results='asis'}
build_toc("04-transform-tables.Rmd")
```

***

## What you should know before you begin {-}

```{block2, type='rmdcaution'}
Tidyverse functions that transform tables do not transform your original table, which is preserved in its original form under its original name. Tidyverse functions instead return a modified copy of the table, which can be passed to the next function in a pipe, or assigned to a name for later use.
```

```{block2, type = "rmdcaution"}
dplyr functions will match column names that appear within their arguments to columns in the data frame that they are manipulating. To avoid this behavior, place `!!` before a name to unquote it, e.g. `!!model`.
```

```{block2, type = "rmdcaution"}
Transforming a table sometimes requires more than one recipe. Why? Because tables are made of multiple data structures that work together: 

1. The table itself is a data frame or tibble.  
1. The columns of the table are vectors.  
1. Some columns may be list-columns, which are lists that contain vectors.  

This matters because one of the [design philosophies of the tidyverse](#what-is-the-tidyverse) is that each function should focus on one task and do it well. Functions that do one task well tend to work with one type of data structure. So to transform a table, you often need to combine a function that manipulates tables with a function that manipulates data vectors and maybe a function that manipulates lists. See [Combine Transform Recipes] for tips.
```

## Arrange rows in ascending order by the values of a column

You want to sort the rows of a data frame in **ascending** order by the values in one or more columns.

#### Solution {-}

```{r}
mpg %>% 
  arrange(cty)
```

#### Discussion {-}

`arrange()` sorts the rows according to the values of the specified column, with the lowest values appearing near the top of the data frame. `arrange()` will use additional columns as tiebreakers to sort within rows that share the same value of the first column.

```{r}
mpg %>% 
  arrange(cty, hwy)
```

## Arrange rows in descending order by the values of a column

You want to sort the rows of a data frame in **descending** order by the values in one or more columns.

#### Solution {-}

```{r}
mpg %>% 
  arrange(desc(cty))
```

#### Discussion {-}

Place `desc()` around a column name to cause `arrange()` to sort by descending values of that column. You can use `desc()` for tie-breaker columns as well.

```{r}
mpg %>% 
  arrange(cty, desc(hwy))
```


## Filter rows with a logical test

You want to filter your table to just the rows that meet a specific condition.

#### Solution {-}

```{r}
mpg %>% 
  filter(model == "jetta")
```

#### Discussion {-}

`filter()` takes a logical test and returns the rows for which the logical test returns `TRUE`. `filter()` will match column names that appear within the logical test to columns in your data frame. 

`filter()` will combine multiple logical tests with an AND operator (`&`):

```{r}
mpg %>% 
  filter(model == "jetta", year == 1999)
```

See `?Comparison` and `?Logic` to learn more about writing logical tests in R.

## Select columns by name

You want to return a "subset" of columns from your data frame by listing the name of each column to return.

#### Solution {-}

```{r}
table1 %>% 
  select(country, year, cases)
```

<!-- I couldn't find table1 anywhere in a search, may want to provide code for people at the top to generate? Or at least cross-reference -->

#### Discussion {-}

`select()` returns a new data frame that includes each column passed to `select()`. Repeat a name to include the column twice.

## Drop columns by name

You want to return a "subset" of columns from your data frame by listing the name of each column to drop.

#### Solution {-}

```{r}
table1 %>% 
  select(-c(population, year))
```

#### Discussion {-}

If you use a `-` before a column name, `select()` will return every column in the data frame except that column. To drop more than one column at a time, group the columns into a vector preceded by `-`. 

## Select a range of columns

You want to return two columns from a data frame as well as every column that appears between them. 

#### Solution {-}

```{r}
table1 %>% 
  select(country:cases)
```

#### Discussion {-}

If you combine two column names with a `:`, `select()` will return both columns and every column that appears between them in the data frame.

## Select columns by integer position

You want to return a "subset" of columns from your data frame by listing the position of each column to return.

#### Solution {-}

```{r}
table1 %>% 
  select(1, 2, 4)
```

#### Discussion {-}

`select()` interprets the whole number _n_ as the _n_th column in the data set. You can combine numbers with `-` and `:` inside of `select()` as well.

## Select columns by start of name

You want to return evey column in your data that begins with a specific string.

#### Solution {-}

```{r}
table1 %>% 
  select(starts_with("c"))
```

## Select columns by end of name

You want to return evey column in your data that ends with a specific string.

#### Solution {-}

```{r}
table1 %>% 
  select(ends_with("tion"))
```

## Select columns by string in name

You want to return evey column in your data whose name contains a specific string or regular expression.

#### Solution {-}

```{r}
table1 %>% 
  select(matches("o.*u"))
```

#### Discussion {-}

`o.*u` is a regular expression that matches an `o` followed by a `u` with any number of characters in between. `country` and `population` are returned because the names `country` and `population` each contain an `o` followed (at any distance) by a `u`.

## Reorder columns

You want to return all of the columns in the original data frame in a new order.

#### Solution {-}

```{r}
table1 %>% 
  select(country, year, population, cases)
```

#### Discussion {-}

Use `select()` to select all of the columns in the new order.

## Reorder columns without naming all columns

You want to reorder some of the columns in the original data frame, but you don't care about the order for other columns, and you may have too many columns to name them each individually.

#### Solution {-}

```{r}
table1 %>% 
  select(country, year, everything())
```

#### Discussion {-}

Use `everything()` within `select()` to select all of the columns in the order they are named: all columns are kept, and no columns are duplicated. Using `everything()` preserves the original ordering of the original (unnamed) columns.

## Rename columns

You want to rename one or more columns in your data frame, retaining the rest.

#### Solution {-}

```{r}
table1 %>% 
  rename(state = country, date = year)
```

#### Discussion {-}

For each column to be renamed, type a new name for the column and set it equal to the old name for the column.

## Return the contents of a column as a vector

You want to return the contents of a single column as a vector, not as a data frame with one column.

#### Solution {-}

```{r}
table1 %>% 
  pull(cases)
```


#### Discussion {-}

`pull()` comes in the dplyr package. You can also pull a column by integer position:

```{r}
table1 %>% 
  pull(3)
```

## Summarise data

You want to compute summary statistics for your data set.

#### Solution {-}

```{r}
table1 %>% 
  summarise(total_cases = sum(cases), max_rate = max(cases/population))
```

#### Discussion {-}

When used by itself, `summarise()` returns a new 1 x n tibble. Each argument name passed to `summarise()` will become a column name in the tibble. Each column will contain the result of the R expression passed to the argument/column name. 

`summarise()` is intended to be used with _summary functions_, which are functions that take a vector of values as input and return a single value as output (e.g `sum()`, `max()`, `mean()`).

`summarize()` is an alias for `summarise()`.

## Summarise data by groups

You want to compute and then compare summary statistics for different groups of data in your data set.

#### Solution {-}

```{r}
table1 %>% 
  group_by(country) %>% 
  summarise(total_cases = sum(cases), max_rate = max(cases/population))
```

#### Discussion {-}

`group_by()` converts your data set into a grouped tibble, which is a tibble subclass that indicates in its attributes which rows belong to which group. 

Where appropriate, tidyverse functions recognize grouped tibbles. The functions: 

1. treat each group as a distinct data set
1. execute their code content separately on each group
1. combine the results in a new data frame that contains a column of group labels (as above).

Pass `group_by()` a column to group by the values of that column (rows that contain the same value will be placed in the same group). Pass `group_by()` two or more column names to group by the distinct combinations of values in those columns.

As a rule, tidyverse functions will return a result that contains the same grouping criteria as the original grouped tibble. `summarise()` is an exception: `summarise()` removes the last column name from the grouping criteria of its result. Because of this, you can call `summarise()` repeatedly to view progressively higher level summaries:

```{r}
table1 %>% 
  group_by(country, year) %>% 
  summarise(total_cases = sum(cases))
```

```{r}
table1 %>% 
  group_by(country, year) %>% 
  summarise(total_cases = sum(cases)) %>% 
  summarise(total_cases = sum(total_cases))
```

```{r}
table1 %>% 
  group_by(country, year) %>% 
  summarise(total_cases = sum(cases)) %>% 
  summarise(total_cases = sum(total_cases)) %>% 
  summarise(total_cases = sum(total_cases))
```

## Mutate data (Add new variables)

You want to compute one or more new variables and add them to your table as columns.

#### Solution {-}

```{r}
table1 %>% 
  mutate(rate = cases/population, percent = rate * 100)
```

#### Discussion {-}

`mutate()` returns a copy of your table that contains additional columns. Each argument name passed to `mutate()` will become the name of a new column. Each new column will contain the result of the R expression passed to the argument/column name. 

`mutate()` will build the columns in the order that you define them. As a result, you may use a column in the column definitions that follow it.

`mutate()` is intended to be used with _vectorized functions_, which are functions that take a vector of values as input and return a new vector of values as output (e.g `abs()`, `round()`, and all of R's math operations).

Use `transmute()` to return only the new columns that `mutate()` would create.

```{r}
table1 %>% 
  transmute(rate = cases/population, percent = rate * 100)
```

  
<!--

## How to combine transform recipes

There are two ways to 


When you transform a data frame, you often work with several functions:

1. A function that transforms the structure of the data frame to add or remove rows or columns. This function transforms a _table_. 

1. A function that transforms the values that appear in the data frame to create new values to place in a new row or column. This function transforms a _data vector_.

1. A function that mediates between 1 and 2 if the values are embedded in a list-column, which is a _list_.

The most concise way to learn how to combine these functions is to learn about the functions in isolation, and to then have the functions call each other as necessary.

### An Example

`mtcars` contains performance data for 32 cars.^[To make the example easier to inspect, I made `mtcars` a tibble and removed the `wt`, `qsec`, `vs`, `am`, `gear` and `carb` columns before running this example.]

```{r echo = FALSE}
mtcars <- mtcars %>% 
  select(-wt, -qsec, -vs, -am, -gear, -carb) %>% 
  as_tibble()
```

```{r}
mtcars
```

To round the `mpg` value for each car, you would need two functions:

1. `round()` which can round the values of the data vector stored in `mtcars$mpg`.^[`mtcars$mpg` is R syntax for the `mpg` column of the `mtcars` data frame.]

```{r}
round(mtcars$mpg)
```

2. `mutate()` which can add the results to a copy of the `mtcars` table. 

```{r}
mtcars %>% 
  mutate(mpg_int = round(mpg))
```

`round()` works with data vectors. `mutate()` works with tables. Together they create the table you want.

### Another Example

`sepals` contains sepal length measurements for three species of flower. The measurements are stored in a list-column named `lengths`.^[I made this data frame with the code `sepals <- iris %>% 
  group_by(Species) %>% 
  summarise(lengths = list(Sepal.Length))`]

```{r echo = FALSE}
sepals <- iris %>% 
  group_by(Species) %>% 
  summarise(lengths = list(Sepal.Length))
```

```{r}
sepals
```

Each cell in `lengths` contains a data vector of 50 sepal lengths.

```{r}
# For example, the first cell of lengths
sepals[[1, 2]]
```

To append the average sepal length for each species, you would need three functions:

1. `mean()` which can compute the average value of a data vector in `lengths`

    ```{r}
    mean(sepals[[1, 2]])
    ```
    
2. `map_dbl()` which can apply `mean()` to each cell of `lengths`, which is a list-column.

    ```{r}
    map_dbl(sepals$lengths, mean)
    ```

3. `mutate()` which can add the results to a copy of the `sepals` table. 

    ```{r}
    sepals %>% 
      mutate(avg_length = map_dbl(lengths, mean))
    ```

`mean()` works with data vectors. `map_dbl()` works with list-columns. `mutate()` works with tables. Together they create the table you want.



## Join data sets

You want to

#### Solution {-}

#### Discussion {-}

## Filter one data set against another

You want to

#### Solution {-}

#### Discussion {-}
-->

