
# Transform Lists

***

```{r echo = FALSE, results='asis'}
build_toc("05-transform-lists.Rmd")
```

***

## What you should know before you begin {-}

```{block2, type='rmdcaution'}
A list is a _recursive vector_: a vector that can contain another vector or list in each of its elements. 
```

```{block2, type='rmdcaution'}
When you speak about R, it is difficult to distinguish between vectors that are lists and vectors that are not lists.  Whenever the difference matters, I'll call the first set of vectors **lists** and the second set of vectors **data vectors**. This chapter focuses on lists.
```

```{block2, type='rmdcaution'}
Lists are one of the most flexible data structures in R. As a result, they are used as a general purpose glue to hold objects together. You will find lists disguised as model objects, data frames, list-columns within data frames, and more.
```

```{block2, type='rmdcaution'}
Data frames are a sub-type of list. Each column of a data frame is an element of the list that the data frame is built around.
```

## Extract an element from a list 

You want to return an element of a list as it is, perhaps to use in a function. You do not want the element to come embedded in a list of length one.

#### Solution {-}

```{r}
# returns the element named x in state.center
state.center %>% 
  pluck("x")
```

#### Discussion {-}

`pluck()` comes in the purrr package and does the equivalent of `[[` subsetting. If you pass `pluck()` a character string, `pluck()` will return the element whose name matches the string. If you pass `pluck()` an integer _n_, `pluck()` will return the _nth_ element.

Pass multiple arguments to `pluck()` to subset multiple times. `pluck()` will subset the result of each argument with the argument that follows, e.g.

```{r eval = FALSE}
library(repurrrsive)
sw_films %>% 
  pluck(7, "title")
```

<!--

## Extract one or more elements from a list as a new list

You want to

#### Solution {-}

#### Discussion {-}

## Apply a function to each element of a list

You want to

#### Solution {-}

#### Discussion {-}
-->